---
title: 'Stats 250: Lec 2 Simulations'
author: "Instructional Team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulation Study 1 - Sample Proportions 

### External Package

The instructional team has created a package (called “stats250package”) that contains a plethora of helpful functions. In addition to this package, we will also be utilizing two other external packages ("ggplot2" and "ggdist"). Load these packages by running the following code chunk:

```{r loadStats250Pacakge}
library(stats250package)
library(ggplot2)
library(ggdist)
```


Read in the data using the code chunk below. 

```{r read_data}
institution <- read.csv("institution_level.csv")

```

Here's a quick preview of the data set and its variables.

```{r preview_institution}
head(institution)
```
Some of the categorical variables include "region", "locale", and "type". From these variables, we might be interested in determining...

- What proportion of colleges and universities are in the Midwest?
- What percent of colleges and universities are in cities?
- What proportion of colleges and universities are private?

Let's focus on the proportion of colleges and universities in the Midwest. Because we have a data set that consists of *all* universities and colleges in the US, we could calculate the *population* proportion of universities and colleges in the US that are in the Midwest. The `table()` function could help us accomplish this!

```{r frequency_table_region}
table(institution$Region)
```

We find that 465 out of the 1842 universities and colleges in the US are in the Midwest. 

```{r population_proportion}
465/1842
```

This value (0.252) represents the *population* proportion of all universities and colleges in the US that are in the Midwest. This value would represent a *parameter*.


### Sample Proportions

Suppose we aren't able to collect information for *every* college and university in the US (for one reason or another). Instead, we can only collect a subset - or *sample* - of universities and colleges. 

The `sampleProportion()` function will draw a random sample from the provided data set and compute the sample proportion.

The function requires the variable of interest (dataset$variable) and the size of the random sample that you wish to take (`n`) and the value of the variable that you wish to compute the sample proportion for (`value`).

Let's calculate the proportion of universities and colleges in the Midwest for a random sample of 10 institutions. The function will return the values of the n observations, along with the computed sample proportion. 

```{r sample_proportion_example}
sampleProportion(institution$Region, n = 10, value = "Midwest")
```

What sample proportion did you get? Is it the same as your instructor? Is it the same as the student next to you? Probably not - because you took a *random sample* of 10 institutions and calculated a sample proportion. 

**This statistic will vary from sample to sample!**

Feel free to run the code chunk above a few times to see how the sample proportion bounces around with each new random sample. 

### Sampling Distribution of pi-hat

We will use the `simProportions()` function to help us visualize how the sample proportion varies from sample to sample. This function will:

1. Take a random sample of size n from the population
2. From this random sample, compute the sample proportion
3. Save the value of the sample proportion computed in Step 2
4. Repeat Steps 1 - 3 many, many times 
5. Create a dot plot of the saved sample proportions

While this function has many arguments, we are only going to focus on two for now:

- `p`: the value of the population proportion
- `n`: the sample size

Using the population proportion of female employees at this company (computed earlier to be 152/400), let's simulate the sampling distribution of phat for repeated random samples of 20 employees. 

```{r sampling_distribution_example}
simProportions(p = 0.25, n = 10)
```


**Question:** In the plot above, what does each dot represent?


## Simulation Study 2 - Sample Means Part 1 
Read in the data using the code chunk below. 

```{r read_data}
tuition <- read.csv("public_tuition.csv")

```

Here's a quick preview of the data set and its variables.

```{r preview_tuition}
head(tuition)
```

Let's create a histogram for the `tuition_out` variable. 

Note: an additional line of code was added to the end of our typical histogram code to add cleaner breaks along the x-axis. Feel free to use this in your own code or ignore it! 

```{r histogram_tuition}
ggplot(data = tuition, aes(x = tuition_out)) + 
  
  geom_histogram(bins = 10, 
                 color = "black", 
                 fill = "white")+
  
  labs(title = "Histogram of Out-of-State Tuition",
       subtitle = "by Stats 250 Instructional Team",
       x = "Out-of-State Tuition ($)",
       y = "Frequency") +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5))
```



How would we describe the distribution of tuition costs for this population? 



Let's calculate the average tuition cost.  

```{r mean_tuition}
mean(tuition$tuition_out)

```

Think About It: Is this value a parameter or a statistic? Why?


### Random Samples and Sample Means

We will use the `sampleMean()` function (from the stats250package) to help us collect a random sample and compute its mean.  

The function takes the following arguments:

- `data`: dataset$variable
- `n`: sample size 

Let's use the function to take a random sample of 10 tuition costs from the tuition data set. 

```{r sampleMean_example_n10}
sampleMean(tuition$tuition_out, n = 10)
```


What sample mean did you get? Is it the same as your instructor? Is it the same as the student next to you? Probably not - because you took a *random sample* of 10 institutions and calculated a sample mean. 

Feel free to run the code chunk above a few times to see how the sample mean bounces around with each new random sample of size 10. 


We can *simulate* the idea of a sampling distribution using the following steps. 

1. Take a random sample of size n from the population of interest
2. With this random sample, compute the sample mean
3. Save the value of the sample mean computed in Step 2
4. Repeat Steps 1 - 3 many, many times 
5. Create a plot of the many, many saved sample means

The `sampleMean()` function generated *one* random sample every time we ran the function. What if we wanted to generate *many, many* random samples - say 500 random samples? We can use the `simMeans()` function!

The function takes the same arguments as the previous function:

- `data`: dataset$variable
- `n`: sample size  

The function will return a plot of 500 sample means calculated from 500 random samples of size n. Note: the value of 500 is not the sample size, but rather the number of sample means we will collect.  

We will start by taking samples of size 10. Although our sample size is 10, the function is:

- taking many, many random samples of size 10
- calculating the sample mean for each random sample of size 10
- saving the value for each of the resulting sample means
- creating a plot of these saved sample means

```{r simMeans_example_n10}
simMeans(tuition$tuition_out, n = 10, xlim=c(0, 50000))
```


## Simulation Study 2 - Sample Means Part 1 


Read in the data using the code chunk below. 

```{r read_data}
enroll <- read.csv("institution_enrollment.csv")
```

Here's a quick preview of the data set and its variables.


```{r preview_enrollment}
head(enroll)
```


Let's create a histogram for the `tuition_out` variable. 

Note: an additional line of code was added to the end of our typical histogram code to add cleaner breaks along the x-axis. Feel free to use this in your own code or ignore it! 

```{r histogram_tuition}
ggplot(data = enroll, aes(x = Enrollment)) + 
  
  geom_histogram(bins = 20, 
                 color = "black", 
                 fill = "white")+
  
  labs(title = "Histogram of Enrollment",
       subtitle = "by Stats 250 Instructional Team",
       x = "Number of students ",
       y = "Frequency") +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5))
```



How would we describe the distribution of enrollments for this population? 



Let's calculate the average tuition cost.  

```{r mean_tuition}
mean(enroll$Enrollment)

```

Think About It: Is this value a parameter or a statistic? Why?


### Random Samples and Sample Means

We will use the `sampleMean()` function (from the stats250package) to help us collect a random sample and compute its mean.  

The function takes the following arguments:

- `data`: dataset$variable
- `n`: sample size 

Let's use the function to take a random sample of 10 tuition costs from the tuition data set. 

```{r sampleMean_example_n10}
sampleMean(enroll$Enrollment, n = 10)
```


What sample mean did you get? Is it the same as your instructor? Is it the same as the student next to you? Probably not - because you took a *random sample* of 10 institutions and calculated a sample mean. 

Feel free to run the code chunk above a few times to see how the sample mean bounces around with each new random sample of size 10. 


We can *simulate* the idea of a sampling distribution using the following steps. 

1. Take a random sample of size n from the population of interest
2. With this random sample, compute the sample mean
3. Save the value of the sample mean computed in Step 2
4. Repeat Steps 1 - 3 many, many times 
5. Create a plot of the many, many saved sample means

The `sampleMean()` function generated *one* random sample every time we ran the function. What if we wanted to generate *many, many* random samples - say 500 random samples? We can use the `simMeans()` function!

The function takes the same arguments as the previous function:

- `data`: dataset$variable
- `n`: sample size  

The function will return a plot of 500 sample means calculated from 500 random samples of size n. Note: the value of 500 is not the sample size, but rather the number of sample means we will collect.  

We will start by taking samples of size 10. Although our sample size is 10, the function is:

- taking many, many random samples of size 10
- calculating the sample mean for each random sample of size 10
- saving the value for each of the resulting sample means
- creating a plot of these saved sample means

```{r simMeans_example_n10}
simMeans(enroll$Enrollment, n = 5, xlim=c(0, 40000))
```




